Visualizing probability distributions across bivariate cyclic temporal granularities

### 2021-04-21 reviews/response.Rmd

Q3. What is the definition of aperiodic linear granularities? On Page 11, Definition 10, $M_i$ are aperiodic linear granularities. If $M_i$ are aperiodic, then is $M$ also an aperiodic linear granularity? What is the relationship between aperiodic linear granularity and aperiodic
cyclic granularity?

A3. Aperiodic linear granularities are those that can not be defined by Definition 6. If $M_i$ is periodic, then $M = \bigcup M_i$ is periodic, but if each $M_i$'s are aperiodic, then $M$ could be periodic or aperiodic. In Definition 10, we want $M$ to be aperiodic. Definition 10 and introduction of Section 3.3 briefly rephrased to improve clarity.

Q4. $w$ in $T_w$ is not well defined for circular granularities in Section 3.2 & 3.5. The definition of $T_w$ uses $T_w$  itself in the formula making it hard to understand. Relevantly, what is the interpretation of $k$ where w = 0; 1; :::; k-1? In my opinion, the number of cycles
should be defined for cyclic granularities, for example, $w_{B,G}(z) = \lfloor z/P(B;G) \rfloor$, which
represent the cycle that the granule is in.

A4. $w$ denotes the index set of $G'$.  $R'$ is the number of granules of $G'$ in each period.$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ does not use the definition of $T_w$. For example, if $B$ is `day` and $G'$ is `month`, then $G'(0) = {B(0)\cup B(2)\cup \dots \cup B(30)}$, $G'(1) = {B(31)\cup B(32)\cup \dots \cup B(58)}$ corresponding to January and February. Then $T_0 = \{0, 1, 2, \dots, 30\}$ and $T_1 = \{31, 32, \dots, 58\}$. If $z$ belongs to $k^{th}$ granule of $G'$, then $z \in T_k$. For example, if $z = 31$ then $z \in T_1$ and thus $k = 1$.
(Should this be added to paper?)

<!-- Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$. The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property. The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$. The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$. Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition 9. -->

@@

Q9. Page 20, Line 47, "for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends" is hard to be justified. By Figure the IQR is given by the height of blue (LV=F) boxes. They are the largest in Quarter 2 and smallest in Quarter 4 for both weekdays and weekends.

A9. removed.

## Other questions of interest.

Q1. Page 6, Definition 5, should $S$ be a contiguous set of integers or not necessary?

@@

A1. Not necessary. 
<!-- Consider an example, where $B$ is the bottom granularity denoting day, $G$ denotes every alternate business days (Monday, Wednesday and Friday) and $H$ denotes weeks. Here, $G \trianglelefteq H$. -->

Q2. Page 6, what is the difference between finer than and group into in Definitions 4 and 5? On Page 7, Line 34, they are used as alternative choices. Can you give an example
that $G \preceq H$ does not imply $G  \trianglelefteq H$, or vice versa? "finer than" is not used a lot in this paper. If the two concepts are similar, why not pick one of them?

A2. If $G \trianglelefteq H$, then every granule of $H$ can be expressed as an union of some set of granules in $G$. If $G \preceq H$, then every granule of $G$ is a subset of some granule of $H$. These are not equivalent.
Consider an example, where $G_1$ denotes "weekend" and $H_1$ denotes "week". Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$. Consider another example, with $G_2$ denotes "days" and $H_2$ denotes "business-week". Then, $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
Further, consider an example with  $H_3$ denoting "public holidays", then $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.    
(Should this be added to paper?)

@@

Q3. Page 16, Line 13, multiple order-up quasi-circular granularities. The equations for this
operation are not trivial to be ignored. Should be added if possible.

A3. Exactly equations (5) and (6) are to be utilized for this operation (with sides interchanged) and hence not added again.

@@

Q4. Can the time granularity concept be extended to continuous time domain?

A4. For continuous time domains, the dependent variable is specified for every value of time $t$, whereas, in applications we generally have the values of the dependent variable for discrete time domain Hence, we will restrict ourselves to explore only discrete time domain.

### 2021-04-22 main.Rmd
comments revisited
made periodic instead of periodical throughout

\begin{definition}\label{def:periodical}
A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{definition}

If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$. Condition (2) in Definition \ref{def:periodical} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$. In this pattern, each granule of $H$ is shifted by $P$ granules of $G$. $P$ is called the \textbf{Period} (@Bettini2000-qk).

For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).

Granularities can also be periodical with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy]. In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

@@

Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodical}. Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant. A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf]. In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduled event. All of these are recurring events, but with non-periodic patterns. Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.

@@

The basic plot choice for our data structure is one that can display distributions. For displaying the distribution of continuous univariate variable, many options are available. Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in [@stryjewski201040]. They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot. Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft]. The less commonly used Letter-Value plots [@Hofmann2017-sg] is a midway between boxplots and density plots. Letter values are order statistics with specific depths. The median ($M$) divides the data set into halves. Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc. are obtained. They are useful for displaying the distributions beyond the quartiles specially for large data,
where boxplots mislabel data points as outliers.
One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices. For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.

>>>>

\begin{definition}\label{def:periodic}
A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{definition}

If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$. Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$. In this pattern, each granule of $H$ is shifted by $P$ granules of $G$. $P$ is called the \textbf{Period} (@Bettini2000-qk).

For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).

Granularities can also be periodic with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy]. In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

@@

Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}. Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
These are events which repeat but there is no reasonably small period within which their behavior remains constant. A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf]. In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduled event. All of these are recurring events, but with non-periodic patterns. Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.

@@

The basic plot choice for our data structure is one that can display distributions. For displaying the distribution of continuous univariate variable, many options are available. Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in [@stryjewski201040]. They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot. Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft]. The less commonly used Letter-Value plots [@Hofmann2017-sg] is a midway between boxplots and density plots. Letter values are order statistics with specific depths. The median ($M$) divides the data set into halves. Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc. are obtained. They are useful for displaying the distributions beyond the quartiles specially for large data,
where boxplots mislabel data points as outliers.
One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices. For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively. In R package `gravitas` [@R-gravitas], boxplots, violin, ridge, letter-value, line and area quantile plots are implemented, but it is potentially possible to use any plots which can display the distribution of the data.
